### 2025-05-28 chapters/background.tex
updated relevant chapters & sections

>>>>

Keystone distinguishes itself from widely adopted commercial Trusted Execution Environments (TEEs) such as ARM TrustZone and Intel Software Guard Extensions (SGX) through its open-source design, extensibility, and architectural modularity. While TrustZone and SGX have demonstrated practical value in securing sensitive computations on commodity hardware, their rigid implementations impose significant limitations on research, customization, and fine-grained hardware/software co-design.

ARM TrustZone implements a two-world model in which the processor switches between a Secure World (TEE) and a Normal World (REE) via secure monitor calls (SMCs). Each world has its own kernel and supports user- and supervisor-level execution. Trusted operating systems such as OP-TEE run in the Secure World and conform to the GlobalPlatform TEE Internal API for developing trusted applications. However, the TrustZone model provisions only a single, statically allocated TEE at boot time, limiting dynamic enclave creation and scalability. Additionally, most interrupts are directed to the Normal World, requiring costly context switches and complicating secure interrupt handling. These constraints, coupled with a relatively large trusted computing base (TCB), hinder the flexibility and security assurances necessary for diverse application scenarios.

Intel SGX adopts a contrasting enclave-based model where enclaves are dynamically created memory regions within user-space processes. These enclaves execute at user level (ring 3) and are protected by hardware-enforced memory encryption. SGX supports secure communication between the host application and the enclave through ECALL and OCALL interfaces, generated and verified using the edger8r tool. While this architecture allows finer-grained protection of sensitive data and logic, SGX lacks supervisor-mode support within enclaves and is tightly coupled to Intel's proprietary hardware platform. Furthermore, enclave memory is pre-allocated at system startup with strict size limitations (e.g., 128MB for SGX v1), restricting its scalability for memory-intensive secure workloads.

Keystone combines and extends the architectural principles of both TrustZone and SGX by enabling dynamic enclave creation, hierarchical privilege separation, and a minimal TCB architecture. Built on the RISC-V instruction set architecture, Keystone features a dedicated Security Monitor (SM) executing in machine mode (M-mode), which governs enclave lifecycle management and access control. Each enclave consists of a user-level application (eapp) and a supervisor-level runtime, allowing greater operational flexibility within the enclave itself. Memory for enclaves is dynamically provisioned by the untrusted operating system and isolated using the RISC-V Physical Memory Protection (PMP) mechanism, ensuring strong spatial and temporal isolation.

To enhance portability and interoperability, Keystone supports a partial implementation of the GlobalPlatform TEE Internal API and introduces keyedger, a secure communication interface similar to SGX's edger8r. Unlike SGX, however, Keystone is designed to be extensible, enabling the integration of advanced security primitives such as secure I/O, remote attestation, and custom scheduling policies. Its open-source nature and hardware-agnostic design position Keystone as a research-friendly and customizable TEE platform, suitable for both academic exploration and real-world deployment scenarios requiring secure, flexible, and verifiable execution environments.

### 2025-05-28 chapters/methodology.tex
updated relevant chapters & sections

\section{Benchmarking tools and metrics}
%\section{Performance metrics}
The performance characterization of Keystone utilized two established synthetic benchmarks tailored for embedded system evaluation: \textit{Dhrystone} and \textit{CoreMark}. Both were selected for their cross-platform compatibility and established use in RISC-V environments.

\textbf{Dhrystone} evaluates integer arithmetic and control flow performance, making it effective for assessing core CPU throughput in environments where floating-point operations are secondary. The RISC-V version, sourced from the \texttt{riscv-tests} repository, was employed to calculate \textit{Dhrystones per second (DMIPS)}. To isolate Keystone’s overhead, measurements were conducted in both native and enclave contexts.

\textbf{CoreMark}, developed by EEMBC, provides a broader performance assessment through representative workloads, including matrix operations, list processing, and state machine evaluations. The RISC-V implementation from the SiFive repository was used to determine \textit{iterations per second (IPS)}. This benchmark offers a deeper view into system performance under more diverse computational tasks than Dhrystone.

Performance variability was assessed by calculating the standard deviation across repeated runs. Metrics collected included:
\begin{itemize}
\item Execution Time
\item DMIPS (Dhrystone) and IPS (CoreMark)
\item Standard Deviation across trials
\end{itemize}

This dual-benchmark approach ensured a comprehensive evaluation of both low-level CPU operations and higher-level application performance within the enclave context.
\section{Parameter Variation Strategy}

>>>>

\section{Benchmarking tools and metrics}

The performance characterization of Keystone utilized two widely adopted synthetic benchmarks: \textit{Dhrystone} and \textit{CoreMark}. Both benchmarks are designed for embedded system evaluation and are well-supported in RISC-V environments. They were chosen to capture the computational characteristics most relevant to TEE deployment, including processor throughput, control flow, and moderate memory usage.

\textit{Dhrystone} primarily evaluates integer arithmetic and control flow performance, making it suitable for measuring core CPU throughput in environments where floating-point operations and I/O are secondary. The RISC-V implementation, sourced from the \texttt{riscv-tests} repository, was used to compute \textit{Dhrystones per second (DMIPS)}. Measurements were taken both in native execution and within a Keystone enclave to quantify the performance overhead introduced by enclave isolation.

\textit{CoreMark}, developed by EEMBC, offers a broader performance profile by incorporating common algorithmic workloads such as list processing, matrix manipulation, and finite-state machine evaluation. While still CPU-bound, CoreMark introduces moderate memory access patterns through its internal data structures. The benchmark was obtained from the SiFive repository and used to calculate \textit{iterations per second (IPS)}.

Performance metrics collected for both benchmarks included:
\begin{itemize}
    \item \textbf{Execution Time:} This is the total elapsed wall-clock time taken for a benchmark to complete execution. It serves as the most direct measure of performance, indicating how long the CPU and memory resources are engaged. Comparing execution times between native and enclave runs highlights the overhead introduced by enclave isolation and security mechanisms.

    \item \textbf{Dhrystones per Second (DMIPS):} Derived from the Dhrystone benchmark, DMIPS represents a normalized measure of integer computing throughput. It quantifies how many Dhrystone operations the CPU can perform per second, serving as a standardized metric for CPU integer performance. A reduction in DMIPS inside the enclave context signals the performance cost of enclave execution.

    \item \textbf{Iterations per Second (IPS):} Obtained from the CoreMark benchmark, IPS measures the number of complete benchmark iterations executed per second. Since CoreMark involves a mixture of algorithmic workloads and memory operations, IPS reflects a broader system performance indicator, capturing both CPU and memory subsystem efficiency under enclave constraints.

    \item \textbf{Standard Deviation:} Calculated across multiple repeated runs of each benchmark, the standard deviation quantifies the variability or consistency of the measured execution times and throughput values. Low standard deviation indicates stable and repeatable performance, while higher values may reveal sensitivity to transient system conditions or measurement noise.

\end{itemize}

\noindent\textbf{Benchmark Characteristics:}
\begin{itemize}
    \item \textit{CPU Intensive:} Both Dhrystone and CoreMark are designed to stress processor logic and control flow. Dhrystone focuses exclusively on integer arithmetic, while CoreMark includes a mix of computational tasks without relying on floating-point operations. These properties make them ideal for evaluating CPU performance in TEE contexts.

    \item \textit{Memory Interaction:} Although not memory-intensive by design, CoreMark exercises moderate memory usage through data structure manipulation, which may expose latency differences due to enclave-related memory isolation. Dhrystone has minimal memory interaction and serves as a purer test of CPU throughput.

    \item \textit{I/O and Storage:} Neither benchmark performs file or persistent storage operations. As such, storage-related TEE APIs (e.g., secure storage or object I/O) are not exercised and are outside the scope of this performance evaluation.
\end{itemize}

This dual-benchmark approach ensures a balanced evaluation of Keystone’s impact on both low-level computational performance and higher-level program behavior in a secure enclave environment.

\section{Parameter Variation Strategy}

### 2025-05-28 main.tex
updated relevant chapters & sections

To run benchmarks, a separate enclave application (eapp) and corresponding host application were developed for each benchmark. The host application initialized and managed the enclave using the Keystone C++ API, while the enclave application contained the benchmark logic and communicated with the host through edge calls. This modular structure allowed each benchmark to be executed both natively (outside the enclave) and securely (within the enclave), enabling accurate measurement of the enclave-related overhead.
To achieve a comparative analysis, two execution modes were considered:
\begin{enumerate}
\item \textit{Native Execution}: Benchmark execution without enclave isolation.
\item \textit{Enclave Execution}: Benchmark execution within the Keystone TEE.
\end{enumerate}

@@

\chapter{Conclusion and Future Work}
\input{chapters/conclusion.tex}

\begin{bibtex-entries}
@TechReport{Kernighan1974,
author = {Brian Kernighan},
title = {Programming in C – A Tutorial},
institution = {Bell Laboratories},
year = {1974}
}
\end{bibtex-entries}
\end{document}

>>>>

To run benchmarks, a separate enclave application (eapp) and corresponding host application were developed for each benchmark. The host application initialized and managed the enclave using the Keystone C++ API, while the enclave application contained the benchmark logic and communicated with the host through edge calls. This modular structure allowed each benchmark to be executed both natively (outside the enclave) and securely (within the enclave), enabling accurate measurement of the enclave-related overhead.

To achieve a comparative analysis, two execution modes were considered:
\begin{enumerate}
\item \textit{Native Execution}: Benchmark execution without enclave isolation.
\item \textit{Enclave Execution}: Benchmark execution within the Keystone TEE.
\end{enumerate}

@@

\chapter{Conclusion}
%\input{chapters/conclusion.tex}

\begin{bibtex-entries}
@TechReport{Kernighan1974,
author = {Brian Kernighan},
title = {Programming in C – A Tutorial},
institution = {Bell Laboratories},
year = {1974}
}

@article{suzaki2021ts,
  title={Ts-perf: General performance measurement of trusted execution environment and rich execution environment on intel sgx, arm trustzone, and risc-v keystone},
  author={Suzaki, Kuniyasu and Nakajima, Kenta and Oi, Tsukasa and Tsukamoto, Akira},
  journal={IEEE Access},
  volume={9},
  pages={133520--133530},
  year={2021},
  publisher={IEEE}
}

\end{bibtex-entries}
\end{document}
